### 2014-10-21 paper/include/background.tex
corrections
started with discussion

The implementation of a DRAM refresh cycle can be made in two different ways, to either affect one row or many. The first is called \textit{RAS-only refresh} (ROR) where a refresh cycle refreshes one targeted row. To target one row, the row's address is put on the bus, and the MC has the responsibility that a row is refreshed within the maximum refresh period. The second is called \textit{CAS before RAS refresh} (CBR). In this scheme, the MC has an internal address counter that hold which global row that will be refreshed. On a refresh cycle, the global row stated by the internal address counter is refreshed and the counter is then incremented. Contrary to ROR, CBR do not need provided addresses on the bus, which results in a lower power consumption and less bus usage. % During CBR, a row in each bank is updated, right? if so. write that in the text.

>>>>

The implementation of a DRAM refresh cycle can be made in two different ways, to either affect one row or many. The first is called \textit{RAS-only refresh} (ROR) where a refresh cycle refreshes one targeted row. To target one row, the row's address is put on the bus, and the MC has the responsibility that a row is refreshed within the maximum refresh period. The second is called \textit{CAS before RAS refresh} (CBR). In this scheme, each rank has an internal address counter that hold which global row that will be refreshed next. On a refresh cycle, a number of global rows are refreshed, beginning with the global row pointed to by the internal address counter, followed by an incrementation of the counter after each refresh of a global row. The number of global rows refreshed is dependent on DRAM device capacity. Contrary to ROR, CBR do not need provided addresses on the bus, which results in a lower power consumption due to less bus usage.

### 2014-10-21 paper/include/reducing_refresh_power.tex
flikker started
reordering of sections
reorder again
started with discussion

\subsubsection*{\textbf{CREAM}}
\label{par:cream}
\input{include/cream}

\subsubsection*{\textbf{Energy Optimization in 3D MPSoCs with Wide-I/O DRAM
Using Temperature Variation Aware Bank-wise Refresh}}
\label{par:tempaware}
\input{include/tempaware}

\subsection{Differences and similarities of the approaches}
\todo[inline]{Drafty text below!}

Differences and similarities for ARTV approaches.

What in the DRAM devices is modified? / Where is data placed.

How much knowledge does the techniques have about the running software?

Smart Refresh:
Employed its solution to 2 GB ram. What about todays sizes? How effective is the technique now? - Only works effective if large parts the memory is accessed, i.e. large part of the memory has to be valid to begin with. Is it so today?

Smart Refresh \& Refrint:
Refrint is somewhat a extension of Smart Refresh. Smart Refresh's features can be laid somewhat equal to Refrint's time-based polices, whereas the data-based polices have no equivalence in Smart Refresh.

They both modify the memory controller, and almost in the same way. But Refrint has no timers that is getting decremented the whole time, and instead copies a global "timer" to the line associated bits upon access. Of course Refrint's data-based polices also adds logic to the memory controller.

Discuss energy savings. Hard to do eDRAM vs DRAM...

>>>>

\subsubsection*{\textbf{Temperature Variation Aware Bank-wise Refresh - TVABR}}
\label{par:tempaware}
\input{include/tempaware}
